A data-intensive assessment of the species-abundance distribution.

### 2015-04-09 defense_pres.tex
More resworking of slides.

\subsection{Challenges of macroecology}
\begin{frame}[t]
\frametitle{Challenges of macroecology}
\begin{itemize}
\item Studies performed with a limited number of large datasets.
\item Lack of identification of pattern generating mechanisms.
\end{itemize}
\note{
Taxonomic and ecosystem limitations in datasets, e.g.North American terrestrial bias.}
\end{frame}

@@

\subsection{Best practices}
\begin{frame}[t]
\frametitle{Best practice recommendations}
\begin{itemize}
\item Test multiple models with consistent statistical approach.
\item Test with multiple taxonomic groups/ecosystems.  
\end{itemize}
\vspace{-7pt}
\begin{center}
\includegraphics[scale=.2]{./sad-data/dissertation/SpatialTaxonomic.png}
\end{center}
\end{frame}

\section{Outline}
\begin{frame}

\end{frame}

\section{Data}
\subsection{Current Data}
\begin{frame}[t]
\frametitle{Current Data}
\vspace{-7pt}
\begin{center}
\includegraphics[scale=.45]{./sad-data/dissertation/ExistingDataMap.png}
\end{center}
\note{
\begin{large}
Major macroecological datasets\\
\end{large}
\begin{itemize}
\item Largely terrestrial
\item Largely North American
\item Many publicly available, some not.
\end{itemize}
~\\
~\\
~\\
~\\
\begin{large}
Lots of data in the literature.\\
\end{large}}
\end{frame}

@@

\section{Species abundance distribution comparisons}
\begin{frame}
\frametitle{SAD Comparisons}
\begin{large}
Most comparisons of the different models:
\end{large}
\begin{itemize}
\item Use a small subset of available models (typically two).
\item Focus on a single ecosystem or taxonomic group
\item Fail to use most appropriate statistical methods. 
\end{itemize}
\end{frame}

@@

\section{Neutral analysis}
\begin{frame}
\frametitle{Neutral Analysis}
Early tests of neutral theory compared the fit of empirical species abundance distributions to the neutral prediction.\\
~\\
Later tests suggested species abundance comparisons were insufficient for a rigorous test of neutrality.\\
~\\
\begin{Large}
\emph{However...}
\end{Large}
\end{frame}

\begin{frame}
\frametitle{Neutral Analysis}
Connolly et al. 2014 simulated neutral communities.\\
\begin{center}
\includegraphics[scale=.25]{./sad-data/dissertation/Connolly2014A.png}
\end{center}
\note{
\begin{small}
\begin{itemize}
\item Compared model fits of a non-neutral distribution (Poisson lognormal) to a neutral distribution (negative binomial distribution).\\
\end{itemize}
\end{small}}
Identified a signal of neutrality.\\
\end{frame}

\begin{frame}
\frametitle{Neutral Analysis}
Connolly et al. 2014 identified non-neutral species abundance distributions in marine communities.\\
\begin{center}
\includegraphics[scale=.25]{./sad-data/dissertation/Connolly2014B.png}
\end{center}
\note{
\begin{small}
\begin{itemize}
\item Compared model fits of a non-neutral distribution (Poisson lognormal) to a neutral distribution (negative binomial distribution).\\
\end{itemize}
\end{small}}
May be a robust method for identifying communities that exhibit non-neutrality.\\
\end{frame}

@@

\begin{frame}[t]
\frametitle{Non-neutral processes}
~\\
Many specific mechanisms, but:
~\\
\begin{itemize}
\item Shape of the abundance distribution due to differences among species.\\
\end{itemize}
\begin{center}
\includegraphics[scale=.2]{./sad-data/dissertation/non-neutral-processes.png}
\end{center}
\end{frame}

@@

\subsection{Conclusions}
\begin{frame}
\frametitle{Neutral analysis}
Difficult to identify a clear winning model.
\begin{itemize}
\item Results consistent with our species abundance distribution model comparisons. 
\item Results different from Connolly et al. 2014.
\begin{itemize}
\item May be biological mechanisms responsible for differences between marine and terrestrial communities.
\note{
\item Non-neutral model outperforms the neutral model in marine systems.
\item Our results suggest marine systems more generally approximated by non-neutral dynamics; terrestrial systems more variable between neutral and non-neutral dynamics.}
\end{itemize}
\end{itemize}
\end{frame}

>>>>

@@

%\subsection{Challenges of macroecology}
%\begin{frame}[t]
%\frametitle{Challenges of macroecology}
%\begin{itemize}
%\item Studies performed with a limited number of large datasets.
%\item Lack of identification of pattern generating mechanisms.
%\end{itemize}
%\note{
%Taxonomic and ecosystem limitations in datasets, e.g.North American terrestrial bias.}
%\end{frame}

%\subsection{Best practices}
%\begin{frame}[t]
%\frametitle{Best practice recommendations}
%\begin{itemize}
%\item Test multiple models with consistent statistical approach.
%\item Test with multiple taxonomic groups/ecosystems.  
%\end{itemize}
%\vspace{-7pt}
%\begin{center}
%\includegraphics[scale=.2]{./sad-data/dissertation/SpatialTaxonomic.png}
%\end{center}
%\end{frame}

\section{Outline}
\begin{frame}
\frametitle{Data-intensive ecology}
\begin{itemize}
\item Leveraging existing ecological data.
\item The species abundance distribution.
\item The statistical approach.
\item The mechanistic approach.
\end{itemize}
\end{frame}

\section{Leveraging existing ecological data}
\subsection{Current Data}
\begin{frame}[t]
\frametitle{Current Data}
\vspace{-7pt}
\begin{center}
\includegraphics[scale=.45]{./sad-data/dissertation/ExistingDataMap.png}
\end{center}
\note{
\begin{large}
Major macroecological datasets\\
\end{large}
\begin{itemize}
\item Largely terrestrial
\item Largely North American
\item Many publicly available, some not.
\end{itemize}
~\\
~\\
~\\
~\\
\begin{large}
Lots of data in the literature.\\
\end{large}}
\end{frame}

@@

\section{SAD comparisons: The statistical approach.}
\begin{frame}
\frametitle{SAD comparisons: The statistical approach.}
\begin{large}
Most comparisons of the different models:
\end{large}
\begin{itemize}
\item Use a small subset of available models (typically two).
\item Focus on a single ecosystem or taxonomic group
\item Fail to use most appropriate statistical methods. 
\end{itemize}
\end{frame}

@@

\section{The mechanistic approach: Neutral analysis.}
\begin{frame}
\frametitle{The mechanistic approach: Neutral analysis.}
Early tests of neutral theory compared the fit of empirical species abundance distributions to the neutral prediction.\\
~\\
Later tests suggested species abundance comparisons were insufficient for a rigorous test of neutrality.\\
~\\
\begin{Large}
\emph{However...}
\end{Large}
\end{frame}

\begin{frame}
\frametitle{Neutral Analysis}
Connolly et al. 2014 simulated neutral communities.\\
\begin{center}
\includegraphics[scale=.25]{./sad-data/dissertation/Connolly2014A.png}
\end{center}
\note{
\begin{small}
\begin{itemize}
\item Compared model fits of a non-neutral distribution (Poisson lognormal) to a neutral distribution (negative binomial distribution).\\
\item Distinct abundance values a measure of power related to the number of species in a community to fit models.
\end{itemize}
\end{small}}
Identified a signal of neutrality.\\
\end{frame}

\begin{frame}
\frametitle{Neutral Analysis}
Connolly et al. 2014 identified non-neutral species abundance distributions in marine communities.\\
\begin{center}
\includegraphics[scale=.25]{./sad-data/dissertation/Connolly2014B.png}
\end{center}
\note{
\begin{small}
\begin{itemize}
\item Compared model fits of a non-neutral distribution (Poisson lognormal) to a neutral distribution (negative binomial distribution).\\
\end{itemize}
\end{small}}
May be a robust method for identifying communities that exhibit non-neutrality.\\
Untested in terrestrial systems.
\end{frame}

@@

\begin{frame}[t]
\frametitle{Non-neutral processes}
\begin{center}
\includegraphics[scale=.4]{./sad-data/dissertation/non-neutral-processes.png}\\
\end{center}
Many specific mechanisms, but:
~\\
\begin{itemize}
\item Shape of the abundance distribution due to differences among species.\\
\end{itemize}
\end{frame}

@@

\subsection{Conclusions}
\begin{frame}
\frametitle{Neutral analysis}
Difficult to identify a clear winning model.
\begin{itemize}
\item Demonstrates the importance of testing with multiple ecosystems.
\end{itemize}
\note{
\begin{itemize}
\item Results consistent with our species abundance distribution model comparisons. 
\item Results different from Connolly et al. 2014.
\begin{itemize}
\item May be biological mechanisms responsible for differences between marine and terrestrial communities.
\note{
\item Non-neutral model outperforms the neutral model in marine systems.
\item Our results suggest marine systems more generally approximated by non-neutral dynamics; terrestrial systems more variable between neutral and non-neutral dynamics.}
\end{itemize}
\end{itemize}}
\end{frame}
